MATH 245 old and unused content

### 2024-04-09 PDEs/math245_hw5.Rmd
finally finished beam problem

---
title: "MATH 245 Homework 5"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-04-07"
output:
  pdf_document:
    extra_dependencies: ["tcolorbox", "xcolor"]
---

@@

@@

Therefore, our solution is:
\begin{tcolorbox}[colback=white, title=Solution]
$$u(t,x)= \sum^{\infty}_{n=1} A_n \,e^{\gamma t} \sin{\left( \frac{n\pi x}{l}\right)}+ \\
\sum^{\infty}_{n=1} B_n \,e^{\zeta t} \sin{\left( \frac{n\pi x}{l}\right)}, \quad \text{where}$$

$$ A_n=\frac{2}{l} \int_{0}^{l}\phi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx-B_n$$
$$B_n=\frac{2\gamma}{(\gamma-\zeta)l} \int_{0}^{l}\phi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx-\frac{2(\gamma-\zeta)}{l} \int_{0}^{l}\psi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx,$$
$$\gamma=\frac{t}{2}(-a-a^2-4b-4\lambda_n c^2)$$
$$\zeta=\frac{t}{2}(-a-a^2-4b-4\lambda_n c^2)$$
$$\lambda_n= \left(\frac{n\pi}{l}\right)^2, \quad n=1,2,3...$$
\end{tcolorbox}

>>>>

---
title: "MATH 245 Homework 5"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-04-08"
output:
  pdf_document:
    extra_dependencies: ["tcolorbox", "xcolor"]
---

@@

If we plug this into the homogeneous Dirichlet BCs for $u$, we find $u(t,0)=T(t)X(0)=0, u(t,L)=T(t)X(L)=0$. In order for both of these to be true, we must have either $T(t)=0$ or $X(0)=X(L)=0$. But if $T(t)=0$, then $u(t,x)=0$ for all $x$, which contradicts our ICs. Thus, we must have $X(0)=X(L)=0$. Similarly, from our other BCs we find that $$X''(0)=X''(L)=0$$.

Plugging our expression $u(t,x)=X(x)T(t)$ into the PDE in (\ref{eqn:beam}), it becomes $$T''(t)X(x)=c^2T(t)X^{(4)}(x)$$
$$\frac{T''(t)}{c^2 T(t)}=\frac{X^{(4)}(x)}{X(x)}=\lambda$$
$$X^{(4)}-\lambda X=0 $$
$$r^4-\lambda =0 $$

### Case 1: Zero eigenvalues, $\lambda=0$

By the PDE, $X^{(4)}(x)=0$ implies that $u_{xxxx}=0$, and thus that $X(x)$ is of the form $Ax^3+Bx^2+Cx+D$. Now plugging in our initial conditions:

$$X(0) = 0 \quad\Longrightarrow\quad D=0$$

$$X''(x)=6Ax+2B \quad\Longrightarrow\quad X''(0)=2B=0 \quad\Longrightarrow\quad B=0$$

$$X''(l)=6Al=0 \quad\Longrightarrow\quad A=0$$

$$X(l)=Cl=0 \quad\Longrightarrow\quad C=0$$

Therefore, we would have $X(x)=0$, so there are no eigenfunctions $X(x)$ that satisfy $X^{(4)}+\lambda X=0$ when $\lambda=0$ and hence no zero eigenvalues.

### Case 2: Negative eigenvalues, $\lambda = -\beta^4 <0$

We will use Green's second identity to show that we cannot have any negative eigenvalues. Start with $X^{(4)}=\lambda X$, multiply by $\bar{X}$, and integrate both sides from 0 to $l$:

\begin{align*}
X^{(4)} &=\lambda X \\
\int_0^l X^{(4)} \bar{X}\,dx &=\lambda \int_0^l  X \bar{X}\,dx \\
&=\lambda \int_0^l  \lvert X \rvert ^2\,dx \end{align*}

Now Green's second identity states that for any $u,v \in C^2 [a,b]$, we have
$$\int_a^bvu'' dx=\int_a^buv'' dx + [vu'-uv']^{x=b}_{x=a}$$

Letting $u= X^{(4)}$ and $v=\bar{X}$,

$$\int_0^l X^{(4)} \bar{X}\; dx= \int_0^l \lvert X^{''} \rvert ^2 \;dx \;+ \left[ \bar{X}X^{(5)}-X^{(4)}\bar{X}' \right]^{x=l}_{x=0}$$
But since $X(0)=X(L)=0$, the boundary terms disappear and we have

$$\int_0^l \lvert X^{''} \rvert ^2 \;dx =\lambda \int_0^l  \lvert X \rvert ^2\,dx$$

$$\lambda = \frac{\int_0^l \lvert X^{''} \rvert ^2 \;dx }{\int_0^l  \lvert X \rvert ^2\,dx} \quad\Longrightarrow\quad \lambda \geq 0$$
Therefore, we have only positive eigenvalues.

### Case 3: Positive eigenvalues, $\lambda = \beta^4 >0$
This case gives us the characteristic equation: $r^4-\beta^4=0$, or $r^2= \pm \beta^2, r=\pm \beta, \pm \beta i$

$$X(x)=Ae^{\beta x}+Be^{-\beta x}+C\sin{(\beta x)}+D\cos{(\beta x)}$$
\begin{equation}\label{eqn:x0}
X(0)=A+B+D=0
\end{equation}

$$X''(x)=\beta^2Ae^{\beta x}+\beta^2Be^{-\beta x}-\beta^2C\sin{(\beta x)}-\beta^2D\cos{(\beta x)}$$
Since we are in a case where we defined $\beta$ to be non-zero, we can divide by it to find that
$$X''(0)=\beta^2A+\beta^2B-\beta^2D=0 \quad\Longrightarrow\quad A+B=D$$
Combining this with (\ref{eqn:x0}), we find that $D=0$.

$$X(L)=Ae^{\beta L}+Be^{-\beta L}+C\sin{(\beta L)}=0$$

$$X''(L)=\beta^2Ae^{\beta L}+\beta^2Be^{-\beta L}-\beta^2C\sin{(\beta L)}=0$$

$$C\sin{(\beta L)}=-C\sin{(\beta L)}, \quad 2C\sin{(\beta L)}=0$$
Either $C=0$ or $\sin{(\beta L)}=0$. Say $C=0$, then $X(0)=A+B=0$ implies $A=-B$. Then the boundary condition gives
$$X(L)=Be^{\beta L}-Be^{-\beta L}=0, \quad B(e^{\beta L}-e^{-\beta L})=0$$

Since $e^{\beta L}$ and $e^{-\beta L}$ are only equal when $L=0$ and we defined $L>0$, we must have $B=A=0$, which would imply that $X(x)=0$, the trivial solution. Therefore, we must have $\sin{(\beta L)}=0$, such that $\beta = \frac{n\pi}{L}$.

Returning to our boundary conditions,

$$X(0)=Ae^0+Be^0+C\sin{0}=0$$

$$A+B=0, \quad A=-B$$

$$X(L)=Ae^{\left(\frac{n\pi}{L}\right) L}-Ae^{-\left(\frac{n\pi}{L}\right) L}+C\sin{\left(\left(\frac{n\pi}{L}\right) L\right)}=0$$

$$X(L)=Ae^{n\pi}-Ae^{-n\pi}+C\sin{(n\pi)}=0$$

$$Ae^{n\pi}=Ae^{-n\pi}$$

But for $n=1,2,3...$, this is only true if $A=0$, which means $B$ is also zero.

Thus, we have eigenvalues of $\lambda_n= \left(\frac{n\pi}{L}\right)^4$ with eigenfunctions 
$$\quad X_n(x)=\sin{\left( \frac{n\pi x}{L}\right)},\quad n=1,2,3...$$

Now we return to the time equation,

$$T''(t)-\lambda c^2 T(t)=0$$

$$T''-\left(\frac{n\pi}{L}\right)^4 c^2 T=0$$

$$r^2-\left(\frac{n\pi}{L}\right)^4 c^2 =0$$

$$T_n(t)=Ae^{\left(\frac{n\pi}{L}\right)^2 ct}+Be^{-\left(\frac{n\pi}{L}\right)^2 ct} $$

$$u_n(t,x)=\left[Ae^{\left(\frac{n\pi}{L}\right)^2 ct}+Be^{-\left(\frac{n\pi}{L}\right)^2 ct} \right]\sin{\left( \frac{n\pi x}{L}\right)},\quad n=1,2,3... $$

Check if this actually satisfies our PDE:
$$u_{tt} =c^2 u_{xxxx} $$

$$u\left(\frac{n\pi}{L}\right)^4c^2 =c^2 \left(\frac{n\pi}{L}\right)^4 u $$

Now we can use the principle of superposition to find our general solution:

$$u(t,x)= \sum_{n=1}^{\infty} \left[Ae^{\left(\frac{n\pi}{L}\right)^2 ct}+Be^{-\left(\frac{n\pi}{L}\right)^2 ct} \right]\sin{\left( \frac{n\pi x}{L}\right)}$$

$$u(0,x)= \phi(x)=\sum_{n=1}^{\infty} \left[A+B \right]\sin{\left( \frac{n\pi x}{L}\right)}$$

$$A+B = \frac{2}{L} \int_{0}^{L}\phi(x)\sin{  \left(\frac{n\pi x}{L}\right) }dx$$

$$u_t(0,x)= \psi(x)=\sum_{n=1}^{\infty} \left(\frac{n\pi}{L}\right)^2 c\left[A-B \right]\sin{\left( \frac{n\pi x}{L}\right)}$$

$$A-B =  \left(\frac{n\pi}{L}\right)^2 \frac{2c}{L} \int_{0}^{L}\psi(x)\sin{  \left(\frac{n\pi x}{L}\right) }dx$$
$$-2B +\frac{2}{L} \int_{0}^{L}\phi(x)\sin{  \left(\frac{n\pi x}{L}\right) }dx=  \left(\frac{n\pi}{L}\right)^2 \frac{2c}{L} \int_{0}^{L}\psi(x)\sin{  \left(\frac{n\pi x}{L}\right) }dx$$
\begin{align*}
B & =\frac{1}{L} \int_{0}^{L}\phi(x)\sin{  \left(\frac{n\pi x}{L}\right) }dx-  \left(\frac{n\pi}{L}\right)^2 \frac{c}{L} \int_{0}^{L}\psi(x)\sin{  \left(\frac{n\pi x}{L}\right) }dx \\
&=\frac{1}{L} \int_{0}^{L} \left[\phi(x)- c\left(\frac{n\pi}{L}\right)^2  \psi(x)\right]\sin{  \left(\frac{n\pi x}{L}\right) }dx \end{align*}

$$A=-B + \frac{2}{L} \int_{0}^{L}\phi(x)\sin{  \left(\frac{n\pi x}{L}\right) }dx$$

Therefore, our solution to (\ref{eqn:beam}) is:
\begin{tcolorbox}[colback=white, title=Solution]
\begin{align*}
& u(t,x)= \sum_{n=1}^{\infty} \left[Ae^{\left(\frac{n\pi}{L}\right)^2 ct}+Be^{-\left(\frac{n\pi}{L}\right)^2 ct} \right]\sin{\left( \frac{n\pi x}{L}\right)}, \quad \text{where} \\ 
& B=\frac{1}{L} \int_{0}^{L} \left[\phi(x)- c\left(\frac{n\pi}{L}\right)^2  \psi(x)\right]\sin{  \left(\frac{n\pi x}{L}\right) }dx\\
& A=\frac{1}{L} \int_{0}^{L} \left[\phi(x)+ c\left(\frac{n\pi}{L}\right)^2  \psi(x)\right]\sin{  \left(\frac{n\pi x}{L}\right) }dx
\end{align*}
\end{tcolorbox}

@@

Therefore, our solution is:
\begin{tcolorbox}[colback=white, title=Solution]
\begin{equation}\label{eqn:teleSol}
u(t,x)= \sum^{\infty}_{n=1} A_n \,e^{\gamma t} \sin{\left( \frac{n\pi x}{l}\right)}+ \\
\sum^{\infty}_{n=1} B_n \,e^{\zeta t} \sin{\left( \frac{n\pi x}{l}\right)},\end{equation}

$$\text{where}  \quad A_n=\frac{2}{l} \int_{0}^{l}\phi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx-B_n$$
$$B_n=\frac{2\gamma}{(\gamma-\zeta)l} \int_{0}^{l}\phi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx-\frac{2(\gamma-\zeta)}{l} \int_{0}^{l}\psi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx,$$
$$\gamma=\frac{t}{2}(-a-a^2-4b-4\lambda_n c^2)$$
$$\zeta=\frac{t}{2}(-a-a^2-4b-4\lambda_n c^2)$$
$$\lambda_n= \left(\frac{n\pi}{l}\right)^2, \quad n=1,2,3...$$
\end{tcolorbox}

Now we use an energy argument to show that the telegraph equation has a unique solution. We start with the given energy equation and differentiate:
\begin{align*}
\frac{d}{dt} E(t) &=\frac{1}{2} \int_{0}^{l} \frac{d}{dt}(u_t^2+c^2u_x^2+bu^2)\;dx \\
                  &=\frac{1}{2} \int_{0}^{l} 2u_tu_{tt}+2c^2u_xu_{xt}+2buu_t)\;dx \\
                  &=\int_{0}^{l} u_tu_{tt}+\int_{0}^{l}c^2u_xu_{xt}+\int_{0}^{l}buu_t \\
                  &= \int_{0}^{l}c^2u_xu_{xt}\; dx = c^2u_xu_t \Bigr\rvert^{l}_0- \int_{0}^{l}c^2u_tu_{xx}\;dx
\end{align*}

The term $c^2u_xu_t \Bigr\rvert^{l}_0$ disappears because of our boundary conditions, so we now have
$$\frac{d}{dt} E(t)=\int_{0}^{l} u_tu_{tt}+buu_t-c^2u_tu_{xx}\;dx = \int_{0}^{l}u_t(u_{tt}+bu-c^2u_{xx})\;dx $$

But our PDE, $u_{tt} + au_t + bu = c^2 u_{xx}$, can be re-written as $u_{tt} + bu - c^2 u_{xx}=-au_t$, which means

$$\frac{d}{dt} E(t)=-\int_{0}^{l}a(u_t^2) \;dx $$

Since $a$ and $u_t^2$ are necessarily non-negative, $\int_{0}^{l}a(u_t^2) \;dx \geq 0$ which means that 
$$\frac{d}{dt} E(t)=-\int_{0}^{l}a(u_t^2) \;dx \leq 0$$.

Since we just showed that the derivative of $E(t)$ is less than or equal to zero for all $t$, we have that $E(t) \leq E(0)$ for all $t>0$.

Let $u$ and $v$ be two solutions to (\ref{eqn:tele}) and define $w=u-v$. Then $w$ satisfies the problem

\begin{equation}\label{eqn:teleW}
    \begin{cases}
     w_{tt} + aw_t + bw = c^2 w_{xx}   & 0<x<l, \quad t>0 \\
     w(0,x) = 0             & 0 \leq x \leq l\\
     w_t(0,x) = 0           & 0 \leq x \leq l\\
     w(t,0)=w(t,l)=0 & t>0
    \end{cases}       
\end{equation}

$$E(0)=\frac{1}{2} \int_{0}^{l}(w_t(0,x))^2+c^2(w_x(0,x))^2+b(w(0,x)^2)\;dx = \frac{1}{2} \int_{0}^{l}(0+0+0)\;dx=0$$

So $E(t) \leq E(0)$ for all $t>0$ but $E(0)=0$ for our solution $w(t,x)$, which means that

$$E(t)=\frac{1}{2} \int_{0}^{l}(w_t^2+c^2w_x^2+bw^2)\;dx \leq 0$$

Since all of the terms under the integrand are non-negative, this is only possible if $w_t=w_x=w=0$ for all $t>0$. Thus, $w=u-v=0 \quad\Rightarrow\quad u=v$, meaning any solution to (\ref{eqn:tele}) is in fact the unique solution.

### 2024-04-09 PDEs/math245_old.Rmd
pde hw 5 prob 5 and 6
finally finished beam problem

>>>>

### HW5

Then the time problem has repeated real roots $r=\frac{-a}{2}$, giving us $T(t)=C_1 e^{-a/2}+C_2te^{-a/2}$.

Now we can return to our solution (\ref{eqn:teleSol}). Since our new IBVP (\ref{eqn:teleW}) has $\phi(x)=\psi(x)=0$, our coefficients $A_n$ and $B_n$ will be zero everywhere, which means $w(t,x)=0$ and therefore $u=v$, so any solution to (\ref{eqn:tele}) is in fact the unique solution.

$$T''+\lambda T=0 $$

$$ r^2+\lambda=0$$

$(r^2-\beta^2)(r^2+\beta^2)=(r+\beta)(r-\beta)(r+\beta i)(r-\beta i)$.

\underline{Case 3:} Negative eigenvalues, $\lambda = -\beta^4 <0$
This case gives us the characteristic equation: $r^4+\beta^4=0$, or $r^2= \pm \beta^2i$.

From $e^{i\theta}=\cos{\theta}+i\sin{\theta}$, we get that $e^{\frac{i\pi}{2}}=i$. Thus, $$r^2=\beta^2e^{\frac{i\pi}{2}}, r^2=\beta^2e^{\frac{-i\pi}{2}}$$
$$r=\beta e^{\frac{i\pi}{4}},\;-\beta e^{\frac{i\pi}{4}},\;\beta e^{\frac{-i\pi}{4}},\;-\beta e^{\frac{-i\pi}{4}} $$
$e^{\frac{\pi}{4}i}=\cos{\frac{\pi}{4}}+i\sin{\frac{\pi}{4}}= \frac{\sqrt{2}}{2}+i\frac{\sqrt{2}}{2}$

$e^{ \frac{-\pi}{4}i}=\cos{\frac{-\pi}{4}}+i\sin{\frac{-\pi}{4}}= \frac{\sqrt{2}}{2}-i\frac{\sqrt{2}}{2}$

$$r = \frac{\beta\sqrt{2}}{2} \pm i\frac{\beta\sqrt{2}}{2}, \frac{-\beta\sqrt{2}}{2} \pm i\frac{\beta\sqrt{2}}{2} $$

Our solutions would then take the form
$$X(x)=e^{\frac{-\beta\sqrt{2}}{2}}\left[A\cos{\left(\frac{\beta\sqrt{2}}{2} x\right)}+B\sin{\left(\frac{\beta\sqrt{2}}{2} x\right)}\right] + e^{\frac{\beta\sqrt{2}}{2}}\left[C\cos{\left(\frac{\beta\sqrt{2}}{2} x\right)}+D\sin{\left(\frac{\beta\sqrt{2}}{2} x\right)}\right]$$

$$X(0)=e^{\frac{-\beta\sqrt{2}}{2}}A+e^{\frac{\beta\sqrt{2}}{2}}C=0 $$
$$A=-e^{\beta\sqrt{2}}C $$
Let $\frac{\beta\sqrt{2}}{2}=\alpha$

$$X(x)=e^{-\alpha}\left[-Ce^{2\alpha}\cos{\left(\alpha x\right)}+B\sin{\left(\alpha x\right)}\right] + e^{\alpha}\left[C\cos{\left(\alpha x\right)}+D\sin{\left(\alpha x\right)}\right]$$

$$X(x)=e^{-\alpha}B\sin{\left(\alpha x\right)}-e^{-\alpha} e^{2\alpha}C \cos{\left(\alpha x\right)} + e^{\alpha}C\cos{\left(\alpha x\right)}+e^{\alpha}D\sin{\left(\alpha x\right)}$$
$$X(x)=e^{-\alpha}B\sin{\left(\alpha x\right)}+e^{\alpha}D\sin{\left(\alpha x\right)}$$

$$X(L)=e^{-\alpha}B\sin{(\alpha L)}+e^{\alpha}D\sin{(\alpha L)}=0$$
$$X''(L)=-\alpha^2e^{-\alpha}B\sin{(\alpha L)}-\alpha^2e^{\alpha}D\sin{(\alpha L)}=0$$

Assume $\sin{(\alpha L)} \neq 0$.

$$e^{-\alpha}B+e^{\alpha}D=0 \quad \Longrightarrow\quad B=-De^{2\alpha}$$
$$X(x)=-De^{2\alpha}e^{-\alpha}\sin{\left(\alpha x\right)}+e^{\alpha}D\sin{\left(\alpha x\right)}$$

Then $X(x)=0$ for all $x$, which is just the trivial solution that doesn't satisfy our ICs. So we must have $\sin{(\alpha L)}=0, \alpha=\frac{\beta\sqrt{2}}{2}=\frac{n \pi}{L}, \beta=\frac{2 n \pi}{\sqrt{2}L}$

<!-- \underline{Case 3:} Positive eigenvalues, $\lambda = \beta^4 >0$ -->

<!-- This case gives us the characteristic equation: $r^4+\beta^4=0$.  -->
<!-- $$(r^2+\beta^2i)(r^2-\beta^2i)=(r+\beta\sqrt{i})(r-\beta\sqrt{i})(r+\beta i^{3/2})(r-\beta i^{3/2})$$ -->
<!-- $$X(x) = A e^{(-1)^{3/4} \beta x} + B e^{-(-1)^{3/4} \beta x} + C e^{(-1)^{1/4} \beta x} + D e^{-(-1)^{1/4} \beta x} $$ -->

<!-- $$X(L) = A e^{(-1)^{3/4} \beta L} + B e^{-(-1)^{3/4} \beta L} + C e^{(-1)^{1/4} \beta L} + D e^{-(-1)^{1/4} \beta L}=0 $$ -->
<!-- $$A+B+C+D=0 $$ -->

<!-- $$X''(x) = -iA\beta^2 e^{(-1)^{3/4} \beta x} -iB\beta^2 e^{-(-1)^{3/4} \beta x} + iC\beta^2 e^{(-1)^{1/4} \beta x} + iD\beta^2 e^{-(-1)^{1/4} \beta x}=0 $$ -->
<!-- $$i\beta^2(-A-B+C+D)=0 $$ -->
<!-- $A+B=C+D$ -->

<!-- $$-A e^{(-1)^{3/4} \beta L} -Be^{-(-1)^{3/4} \beta L} + C e^{(-1)^{1/4} \beta L} + D e^{-(-1)^{1/4} \beta L}=0 $$ -->

\underline{Case 2:} Negative eigenvalues, $\lambda = -\beta^2 <0$
This case gives us the characteristic equation: $r^2-\beta^2=0$. Since our characteristic equation has distinct real roots $r= \pm \beta$, our solutions take the form 
$$T(t)=Ae^{\beta t}+Be^{-\beta t}$$

$\beta A - \beta B = \psi(x)$
$A+B=\phi(x)$
